---
title: "Assignment 5"
author: "Kristin Gmunder"
date: "9/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercises
1. Create a list with four elements:
    a. the character string `"Normal"` to represent the distribution
    b. a named vector of the population parameters
    c. a vector of 15 samples from a normal distribution with mean 2 and variance 5
    d. the components of a histogram of these samples
```{R}
xN_ls <- list(
  distName = "Normal", 
  parameters = c(mean = 2, variance = 5), 
  sample = rnorm(n = 15, mean = 2, sd = sqrt(5))
  )
xN_ls[["histogram"]] <- hist(xN_ls[[3]])
```
    
2. Print the distribution name and parameters (at the same time), and plot the saved histogram.
```{R}
xN_ls[1:2]
plot(xN_ls[[4]])
```

3. Print the summary and variance of this sample.
```{R}
summary(xN_ls[[3]])
var(xN_ls[[3]])
```

4. Test the hypothesis that these observations came from the distribution specified. Store the $p$-value, test statistic, critical value, and decision (reject or fail to reject) as new elements of the list.

$H_0: \mu = \mu_0;\quad H_A: \mu \ne \mu_0.$

```{r}
ttest_N <- t.test(
  x = xN_ls$sample,
  alternative = "two.sided",
  mu = xN_ls[[2]][[1]]
)
ttest_N$statistic
ttest_N$p.value
critVal_N <- qt(0.975, df = 15 - 1)
abs(ttest_N$statistic) > critVal_N
xN_ls$pVal <- ttest_N$p.value
xN_ls$testStat <- ttest_N$statistic
xN_ls$critVal <- critVal_N
xN_ls$decision <- "Fail to Reject the Hypothesis that the True Mean Equals 2"
xN_ls
```

5. Repeat this for the rest of the distributions we inspected in Assignment 4. Did you reject the null hypothesis for any of the distributions? If so, which ones. If so, why do you think you rejected the null hypothesis for these distributions? If not, why not?

####Student's T Distribution:
$H_0: \mu = \mu_0;\quad H_A: \mu \ne \mu_0.$
```{R}
set.seed(1234)
xT_ls <- list(
  distName = "Student's T-distribution", 
  parameters = c(mean = 0, df = 4), 
  sample = rt(n = 15, df = 4)
  )
xT_ls[["histogram"]] <- hist(xT_ls[[3]])
```

```{r}
ttest_T <- t.test(
  x = xT_ls$sample,
  alternative = "two.sided",
  mu = xT_ls[[2]][[1]]
)
ttest_T$statistic
ttest_T$p.value
critVal_T <- qt(0.975, df = 15 - 1)
abs(ttest_T$statistic) > critVal_T
xT_ls$pVal <- ttest_T$p.value
xT_ls$testStat <- ttest_T$statistic
xT_ls$critVal <- critVal_T
xT_ls$decision <- "Fail to Reject the Hypothesis that the True Mean Equals 2"
xT_ls
```

####Chi-Squared Distribution:
$H_0: \mu = \mu_0;\quad H_A: \mu \ne \mu_0.$
```{r}
set.seed(1234)
xC_ls <- list(
  dist   = "Chi-Squared",
  params = c(df = 2),
  sample = rchisq(n = 15, df = 2)
)
xC_ls[["histogram"]] <- hist(xC_ls[[3]])
```

```{R}
hist(xC_ls$sample)
ttest_C <- t.test(
  x = xC_ls$sample,
  alternative = "two.sided",
  mu = xC_ls$params
)
ttest_C$statistic
ttest_C$p.value
critVal_C <- qt(0.975, df = 15 - 1)
abs(ttest_C$statistic) > critVal_C
xC_ls$pVal <- ttest_C$p.value
xC_ls$testStat <- ttest_C$statistic
xC_ls$critVal <- critVal_C
xC_ls$decision <- "Fail to Reject the Hypothesis that the True Mean Equals 2"
xC_ls
```

####F Distribution:
$H_0: \mu = \mu_0;\quad H_A: \mu \ne \mu_0.$
```{R}
set.seed(1234)
xF_ls <- list(
  distName = "F distribution", 
  parameters = c(mean = 1.2, df1 = 90, df2 = 12), 
  sample = rf(n = 15, df1 = 90, df2 = 12)
  )
xF_ls[["histogram"]] <- hist(xF_ls[[3]])
```

```{r}
ttest_F <- t.test(
  x = xF_ls$sample,
  alternative = "two.sided",
  mu = xF_ls[[2]][[1]]
)
ttest_F$statistic
ttest_F$p.value
critVal_F <- qt(0.975, df = 15 - 1)
abs(ttest_F$statistic) > critVal_F
xF_ls$pVal <- ttest_F$p.value
xF_ls$testStat <- ttest_F$statistic
xF_ls$critVal <- critVal_F
xF_ls$decision <- "Fail to Reject the Hypothesis that the True Mean Equals 1.2"
xF_ls
```

####Binomial Distribution:
$H_0: \mu = \mu_0;\quad H_A: \mu \ne \mu_0.$
```{R}
set.seed(1234)
xB_ls <- list(
  distName = "Binomial distribution", 
  parameters = c(mean = 6, size = 9, prob = (2/3)), 
  sample = rbinom(n = 15, size = 9, prob = (2/3))
  )
xB_ls[["histogram"]] <- hist(xB_ls[[3]])
```

```{r}
ttest_B <- t.test(
  x = xB_ls$sample,
  alternative = "two.sided",
  mu = xB_ls[[2]][[1]]
)
ttest_B$statistic
ttest_B$p.value
critVal_B <- qt(0.975, df = 15 - 1)
abs(ttest_B$statistic) > critVal_B
xB_ls$pVal <- ttest_B$p.value
xB_ls$testStat <- ttest_B$statistic
xB_ls$critVal <- critVal_B
xB_ls$decision <- "Fail to Reject the Hypothesis that the True Mean Equals 6"
xB_ls
```

####Negative Binomial Distribution:
$H_0: \mu = \mu_0;\quad H_A: \mu \ne \mu_0.$
```{R}
set.seed(1234)
xNB_ls <- list(
  distName = "Negative Binomial distribution", 
  parameters = c(mean = 5, size = 5, prob = 0.5), 
  sample = rnbinom(n = 15, size = 5, prob = 0.5)
  )
xNB_ls[["histogram"]] <- hist(xNB_ls[[3]])
```

```{r}
ttest_NB <- t.test(
  x = xNB_ls$sample,
  alternative = "two.sided",
  mu = xNB_ls[[2]][[1]]
)
ttest_NB$statistic
ttest_NB$p.value
critVal_NB <- qt(0.975, df = 15 - 1)
abs(ttest_NB$statistic) > critVal_NB
xNB_ls$pVal <- ttest_NB$p.value
xNB_ls$testStat <- ttest_NB$statistic
xNB_ls$critVal <- critVal_NB
xNB_ls$decision <- "Fail to Reject the Hypothesis that the True Mean Equals 5"
xNB_ls
```

####Poisson Distribution:
$H_0: \mu = \mu_0;\quad H_A: \mu \ne \mu_0.$
```{R}
set.seed(1234)
xP_ls <- list(
  distName = "Poisson distribution", 
  parameters = c(lambda = 3), 
  sample = rpois(n = 15, lambda = 3)
  )
xP_ls[["histogram"]] <- hist(xP_ls[[3]])
```

```{r}
ttest_P <- t.test(
  x = xP_ls$sample,
  alternative = "two.sided",
  mu = xP_ls[[2]][[1]]
)
ttest_P$statistic
ttest_P$p.value
critVal_P <- qt(0.975, df = 15 - 1)
abs(ttest_P$statistic) > critVal_P
xP_ls$pVal <- ttest_P$p.value
xP_ls$testStat <- ttest_P$statistic
xP_ls$critVal <- critVal_P
xP_ls$decision <- "Fail to Reject the Hypothesis that the True Mean Equals 3"
xP_ls
```



