---
title: "Assignment 1"
author: "Kristin Gmunder"
date: "8/28/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Exercise Set 1

1. Create an object called "myAge" that holds your age *in months*. Hint: the
multiplication operator in R is "*".
```{r}
myAge <- 21*12+2
```
2. Using the object "myAge", approximate the number of days you've been alive. Use 30.44 days per month. Check this against your age in days on Wolfram Alpha. How many days off was your estimate?
```{r}
myAge*30.44
```
My estimate was exactly the same as in Wolfram Alpha.

3. Using the object "myAge", find your decimal age. Hint: the division operator is "/".
```{r}
(myAge*30.44)/365.25
```

4. Copy and Paste MyAge into the console and press ENTER. What happens? Why?
I got an output of "Error: object 'MyAge' not found," because R is case sensitive.  Therefore, R does not recognize the variable I was using (myAge) as the same as MyAge.


## Exercise Set 2

1. Five people start the newest diet craze. Their weights (in kg) before and after were [78 72 78 79 105] and [67 65 79 70 93], respectively. Store these two vectors as "wtBefore" and "wtAfter", respectively. Find the total weight lost, the average weight lost, and the weight lost per person. Hint: see the ?sum and ?mean functions.
```{R}
wtBefore <- c(78, 72, 78, 79, 105)
wtAfter <- c(67, 65, 79, 70, 93)
totalLost <- wtBefore-wtAfter
totalLost
avgWeight <- sum(totalLost[1:5], na.rm = FALSE) / 5
avgWeight
totalLost[1]
totalLost[2]
totalLost[3]
totalLost[4]
totalLost[5]
```

2. Create the following vectors using the rep() and seq() functions:
    a) the first non-negative integer powers of 2
don't understand question
    
    b) 1, 1.5, 2, 2.5, ..., 12
```{R}
seq(from = 1, to = 12, by = 0.5)
```
    c) 1, 8, 27, 64, ..., 1000
    d) 1, -1/2, 1/3, -1/4, ..., -1/100

    e) 1, 0, 3, 0, 5, 0, 7, 0, ..., 0, 49
    f) 1, 3, 6, 10, 15, ..., 210 (Hint: ?cumsum)
    g) 1, 2, 2, 3, 3, 3, ..., 10, 10, 10, 10, 10, 10, 10, 10, 10, 10 (Hint: ?rep)
3. (ADVANCED) The jth term of the Taylor-MacLaurin series for the natural logarithm is   $$
\log(1 - x) := - \lim_{n \to \infty} \sum \limits_{j = 1}^n \frac{x ^ j}{j}.
  $$

Find the residuals of this approximation for n = 5, 10, 50, 100 around x = 0.25





## Exercise Set 3

1. Create a vector of the months of the year and store it in an object called "mons".
2. You would like to discard the first three months, so you type "mons[-1:3]". What happened? Why should you type instead?
3. Find a method to use the "length()" function to find the last four entries of *any* atomic vector. Apply this to find the last four months of the year.
4. Logically subset the uppercase vector of English letters in the following ways:
    a) Type LETTERS and press ENTER. What is your guess to find the lower case letters?
    b) Create a vector of the first 12 upper-case English letters
    c) Create a vector of the 1st, 3rd, 5th, 7th, ..., 25th English letters
    d) Create a vector of the consonants. (Hint: it may be easier to find the letters that are *not* one of the vowels)
