---
title: "Statistical Distributions"
author: "Kristin Gmunder"
date: "8/30/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Random variables in `R` are shown on <https://stat.ethz.ch/R-manual/R-devel/library/stats/html/Distributions.html>

# Discrete Random Variables

##These are the distributions:

We have a common examples of discrete random variables.  Each distribution also has a corresponding implementation in `R`.

1. The Bernoulli Distribution: <https://en.wikipedia.org/wiki/Bernoulli_distribution>
```{R}
#Random Bernoulli with a fair coin
rbinom(n=1, size=1, prob=0.5)

#Weighted coin
rbinom(n=1, size=1, prob=0.75)
```
2. The Binomial Distribution:
<https://en.wikipedia.org/wiki/Binomial_distribution>
```{R}
#Flip 10 fair coins and record the results
rbinom(n=10, size=1, prob=0.5)

#Flip 10 fair coins and record the number of successes
rbinom(n=1, size=10, prob=0.5)

```

3. The Poisson Distribution:
<https://en.wikipedia.org/wiki/Poisson_distribution>
```{R}
#How many horse kicks in each group of 10 soldiers? We expect 2 horse kicks
rpois(n=10, lambda=2)

#Now the horses are angry...
rpois(n=10, lambda=4)
```
## Moments of these Distributions
The common moments we would care about are the *first* and *second* moments, from which we calculate the mean and variance, respectively.

### Binomial
```{r}
# Random draw of five fair coins

#Mean
xBinom <- rbinom(n=5, size=1, prob=0.5)
xBinom
mean(xBinom)

#Variance
var(xBinom)

# Random draw of 100 fair coins

#Mean
xBinom2 <- rbinom(n=100, size=1, prob=0.5)
xBinom2
mean(xBinom2)

#Variance
var(xBinom2)

```


# Continuous Random Variables
