---
title: "Assignment 3"
author: "Kristin Gmunder"
date: "10/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R for Data Science Assignment

```{R}
df <- tibble::tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

df$a <- (df$a - min(df$a, na.rm = TRUE)) / 
  (max(df$a, na.rm = TRUE) - min(df$a, na.rm = TRUE))
df$b <- (df$b - min(df$b, na.rm = TRUE)) / 
  (max(df$b, na.rm = TRUE) - min(df$a, na.rm = TRUE))
df$c <- (df$c - min(df$c, na.rm = TRUE)) / 
  (max(df$c, na.rm = TRUE) - min(df$c, na.rm = TRUE))
df$d <- (df$d - min(df$d, na.rm = TRUE)) / 
  (max(df$d, na.rm = TRUE) - min(df$d, na.rm = TRUE))

x <- df$a
(x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))

rng <- range(x, na.rm = TRUE)
(x - rng[1]) / (rng[2] - rng[1])

rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
rescale01(c(0, 5, 10))
rescale01(c(-10, 0, 10))
rescale01(c(1, 2, 3, NA, 5))

df$a <- rescale01(df$a)
df$b <- rescale01(df$b)
df$c <- rescale01(df$c)
df$d <- rescale01(df$d)

x <- c(1:10, Inf)
rescale01(x)

rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE, finite = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
rescale01(x)
```

##19.2.1 Practice

1. na.rm is not a parameter of rescale01(), because we are inputting only x into the function and na.rm here serves as an argument in the rng function to let R know that we want to keep NA values while completing this function.  If x contained a single missing value, and na.rm was FALSE, the function would return an error and NA as it does not know what to do, because it does not want to keep NA values.

2. 
```{r}
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
  y = -Inf <- 0
  y = Inf <- 11
}
```

3. 
```{r}
whereisna <- function(x) {
  mean(is.na(x))
}  

otherfunction <- function(x, y) {
  x / sum(x, na.rm = y)
}  

calculation <- function(x, y) {
  sd(x, na.rm = y) / mean(x, na.rm = y)
}  
```

4. 
```{r}
variance <- function(x) {
    n <- length(x)
    m <- mean(x)
    (1/(n - 1)) * sum((x - m)^2)
}

skewness <- function(x) {
    n <- length(x)
    v <- var(x)
    m <- mean(x)
    third.moment <- (1/(n - 2)) * sum((x - m)^3)
    third.moment/(var(x)^(3/2))
}
```    

5.
```{r}
both_na <- function(x, y) {
sum(is.na(x), is.na(y))  
}
```

6.
```{r}
is_directory <- function(x) file.info(x)$isdir
is_readable <- function(x) file.access(x, 4) == 0
```
is_directory checks if the file "x" is a directory.  is_readable checks if it can open and use file "x" (if it is "readable").  These functions are useful because their names tell you what they're doing.

7. 
repetitive <- function() {
  "Little bunny Foo Foo
  Hopping through the forest
  Scooping up the field mice
  And bopping them on the head"
}

chances <- function(chancenumber) {
 "Down came the Good Fairy, and she said
  'Little bunny Foo Foo
  I don't want to see you
  Scooping up the field mice
  And bopping them on the head.
  I'll give you "(chancenumber) " chances,
  And if you don't stop, I'll turn you into a GOON!'"
}

dayrepeat <- function() {
  "And the next day..."
}

end <- function() {
  "'I gave you three chances, and you didn't stop; so....'
  She turned him into a GOON!"
}

repetitive()
chances(3)
dayrepeat()
repetitive()
chances(2)
dayrepeat()
repetitive()
chances(1)
dayrepeat()
repetitive()
end()


f()
my_awesome_function()
impute_missing()
collapse_years()

col_mins <- function(x, y) {}
rowMaxes <- function(y, x) {}

input_select()
input_checkbox()
input_text()
select_input()
checkbox_input()
text_input()

T <- FALSE
c <- 10
mean <- function(x) sum(x)

##19.3.1 Exercises
1. 
```{r}
f1 <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}
f2 <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
f3 <- function(x, y) {
  rep(y, length.out = length(x))
}
```
f1 checks to see if the prefix is in the string. A better name would be check_string.  f2 gets rid of the last entry in x.  A better name would be remove_last.  f3, replicates y, x amount of times.  A better name would b y_x_times.

2. A better name for both_na in number 5 from above would be total_num_NA.  Because it finds the total number of NA entries in both functions.

3. rnorm has arguments of mean, sd, and n (number of observations).  MASS:: mvrnorm has arguments mu, sigma, and n (number of samples).  Perhaps, to make them more consistent their arguments should have the same names.

4. For a family of functions that do similar things, it is best that they have the same prefix.  This is because autocomplete allows us to type the prefix and see the whole family of functions.  Thus norm_r() and norm_d() are better.  On the other hand, the underscore may be irritating to type, so perhaps rnorm and dnorm could be better.

```{r}
has_name <- function(x) {
  nms <- names(x)
  if (is.null(nms)) {
    rep(FALSE, length(x))
  } else {
    !is.na(nms) & nms != ""
  }
}
```
if (c(TRUE, FALSE)) {}
if (NA) {}

```{r}
identical(0L, 0)

x <- sqrt(2) ^ 2
x
x == 2
x - 2
```
function(x, y, op) {
  switch(op,
    plus = x + y,
    minus = x - y,
    times = x * y,
    divide = x / y,
    stop("Unknown op!")
  )
}

if (y < 0 && debug) {
  message("Y is negative")
}

if (y == 0) {
  log(x)
} else {
  y ^ x
}

if (y < 0 && debug)
message("Y is negative")

if (y == 0) {
  log(x)
} 
else {
  y ^ x
}

```{r}
y <- 10
x <- if (y < 20) "Too low" else "Too high"

if (y < 20) {
  x <- "Too low" 
} else {
  x <- "Too high"
}
```

##19.4.4 Exercises
1. For an "if" statement, the code will only execute else if "if" is false.  For an "ifelse", the code will only execute if the "ifelse" is not done and the "ifelse" is true.

2. if (time < 12) {
  "good morning"
  } 
  else if (time < 17) {
    "good afternoon"
  } 
  else {
   "good evening"
  }

3. 
fizzbuzz <- function(x) {
  if (x / 3 % == 0) {
    message("fizz")
  } 
  if (x / 5 % == 0) {
    message("buzz")
  }  
  else {
    x
  }  
}


4. 

5. switch() returns the entry corresponding to that number.

6. 
```{r}
switch("a", 
  a = ,
  b = "ab",
  c = ,
  d = "cd"
) 
```
switch() returns the fist entry that is not missing for the argument that you give it.  So for "a", it returns "ab", because a = " ".

```{r}
mean_ci <- function(x, conf = 0.95) {
  se <- sd(x) / sqrt(length(x))
  alpha <- 1 - conf
  mean(x) + se * qnorm(c(alpha / 2, 1 - alpha / 2))
}

x <- runif(100)
mean_ci(x)
mean_ci(x, conf = 0.99)

mean(1:10, na.rm = TRUE)
```
average <- mean(feet / 12 + inches, na.rm = TRUE)
```{r}
wt_mean <- function(x, w) {
  sum(x * w) / sum(w)
}
wt_var <- function(x, w) {
  mu <- wt_mean(x, w)
  sum(w * (x - mu) ^ 2) / sum(w)
}
wt_sd <- function(x, w) {
  sqrt(wt_var(x, w))
}

wt_mean(1:6, 1:3)

wt_mean <- function(x, w) {
  if (length(x) != length(w)) {
    stop("`x` and `w` must be the same length", call. = FALSE)
  }
  sum(w * x) / sum(w)
}

wt_mean <- function(x, w, na.rm = FALSE) {
  if (!is.logical(na.rm)) {
    stop("`na.rm` must be logical")
  }
  if (length(na.rm) != 1) {
    stop("`na.rm` must be length 1")
  }
  if (length(x) != length(w)) {
    stop("`x` and `w` must be the same length", call. = FALSE)
  }
  
  if (na.rm) {
    miss <- is.na(x) | is.na(w)
    x <- x[!miss]
    w <- w[!miss]
  }
  sum(w * x) / sum(w)
}

t_mean <- function(x, w, na.rm = FALSE) {
  stopifnot(is.logical(na.rm), length(na.rm) == 1)
  stopifnot(length(x) == length(w))
  
  if (na.rm) {
    miss <- is.na(x) | is.na(w)
    x <- x[!miss]
    w <- w[!miss]
  }
  sum(w * x) / sum(w)
}
```
wt_mean(1:6, 6:1, na.rm = "foo")

sum(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)

stringr::str_c("a", "b", "c", "d", "e", "f")

commas <- function(...) stringr::str_c(..., collapse = ", ")
commas(letters[1:10])

rule <- function(..., pad = "-") {
  title <- paste0(...)
  width <- getOption("width") - nchar(title) - 5
  cat(title, " ", stringr::str_dup(pad, width), "\n", sep = "")
}
rule("Important output")

x <- c(1, 2)
sum(x, na.mr = TRUE)

##19.5.5 Exercises
1. commas(letters, collapse = "-") gives an error message "formal argument "collapse" matched by multiple actual arguments".

2. "pad" can only take one character, because this is how it is defined.

3. The trim function cuts out observations from each end of x before computing the mean.  This could be used when there are outliers.

4. The defaults are reffering to different ways to calculate the correlation coefficient.  The default is the pearson correlation coefficient.


show_missings <- function(df) {
  n <- sum(is.na(df))
  cat("Missing values: ", n, "\n", sep = "")
  
  invisible(df)
}

show_missings(mtcars)
x <- show_missings(mtcars) 
class(x)
dim(x)

mtcars %>% 
  show_missings() %>% 
  mutate(mpg = ifelse(mpg < 20, NA, mpg)) %>% 
  show_missings() 

f <- function(x) {
  x + y
} 

y <- 100
f(10)

y <- 1000
f(10)

`+` <- function(x, y) {
  if (runif(1) < 0.1) {
    sum(x, y)
  } else {
    sum(x, y) * 1.1
  }
}
table(replicate(1000, 1 + 2))

rm(`+`)







