---
title: "Assignment1"
author: "Kristin Gmunder"
date: "10/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Bootstrap Assignment 1: Due Sunday, 7 October at 23:59PM
*For help with Rmarkdown for reports, see this [white paper](http://www.stat.cmu.edu/~cshalizi/rmarkdown/#math-in-r-markdown) from Carnegie Mellon University's Department of Statistics and Data Science.*

### Part 2
Create an Rmarkdown file called "Assignment1.Rmd". In this report:

1. Work through this lesson on functions: https://swcarpentry.github.io/r-novice-inflammation/02-func-R/

```{r}
fahrenheit_to_kelvin <- function(temp_F) {
  temp_K <- ((temp_F - 32) * (5 / 9)) + 273.15
  return(temp_K)
}

fahrenheit_to_kelvin(32)
fahrenheit_to_kelvin(212)

kelvin_to_celsius <- function(temp_K) {
  temp_C <- temp_K - 273.15
  return(temp_C)
}

kelvin_to_celsius(0)

fahrenheit_to_celsius <- function(temp_F) {
  temp_K <- fahrenheit_to_kelvin(temp_F)
  temp_C <- kelvin_to_celsius(temp_K)
  return(temp_C)
}

fahrenheit_to_celsius(32.0)

kelvin_to_celsius(fahrenheit_to_kelvin(32.0))

best_practice <- c("Write", "programs", "for", "people", "not", "computers")
asterisk <- "***"  # R interprets a variable with a single value as a vector
                   # with one element.

fence <- function(original, wrapper) {
  answer <- c(wrapper, original, wrapper)
  return(answer)
}

fence(best_practice, asterisk)

outside <- function(v) {
  first <- v[1]
   last <- v[length(v)]
   answer <- c(first, last)
   return(answer)
}

dry_principle <- c("Don't", "repeat", "yourself", "or", "others")
outside(dry_principle)

input_1 <- 20
mySum <- function(input_1, input_2 = 10) {
  output <- input_1 + input_2
  return(output)
}

#mySum(input_1 = 1, 3) gives an output of 4

#If mySum(3) returns 13, why does mySum(input_2 = 3) return an error?
#mySum(input_2=3) returns an error, because the function does not have a value for the argument, input_1, to operate on.
center <- function(data, desired) {
  new_data <- (data - mean(data)) + desired
  return(new_data)
}

z <- c(0, 0, 0, 0)
z

center(z, 3)

dat <- read.csv(file = "/Users/KristinGmunder/Documents/GitHub/BST650_introR/data/inflammation-01.csv", header = FALSE)
centered <- center(dat[, 4], 0)
head(centered)

min(dat[, 4])

mean(dat[, 4])

max(dat[, 4])

min(centered)

mean(centered)

max(centered)

sd(dat[, 4])

sd(centered)

sd(dat[, 4]) - sd(centered)

all.equal(sd(dat[, 4]), sd(centered))

center <- function(data, desired) {
  # return a new vector containing the original data centered around the
  # desired value.
  # Example: center(c(1, 2, 3), 0) => c(-1, 0, 1)
  new_data <- (data - mean(data)) + desired
  return(new_data)
}

analyze <- function(filename) {
  # Plots the average, min, and max inflammation over time.
  # Input is character string of a csv file.
  dat <- read.csv(file = filename, header = FALSE)
  avg_day_inflammation <- apply(dat, 2, mean)
  plot(avg_day_inflammation)
  max_day_inflammation <- apply(dat, 2, max)
  plot(max_day_inflammation)
  min_day_inflammation <- apply(dat, 2, min)
  plot(min_day_inflammation)
}

rescale <- function(v) {
  # Rescales a vector, v, to lie in the range 0 to 1.
  L <- min(v)
  H <- max(v)
  result <- (v - L) / (H - L)
  return(result)
}

dat <- read.csv("/Users/KristinGmunder/Documents/GitHub/BST650_introR/data/inflammation-01.csv", FALSE)

center <- function(data, desired = 0) {
  # return a new vector containing the original data centered around the
  # desired value (0 by default).
  # Example: center(c(1, 2, 3), 0) => c(-1, 0, 1)
  new_data <- (data - mean(data)) + desired
  return(new_data)
}

test_data <- c(0, 0, 0, 0)
center(test_data, 3)

more_data <- 5 + test_data
more_data

center(more_data)

display <- function(a = 1, b = 2, c = 3) {
  result <- c(a, b, c)
  names(result) <- c("a", "b", "c")  # This names each element of the vector
  return(result)
}

# no arguments
display()

# one argument
display(55)

# two arguments
display(55, 66)

# three arguments
display(55, 66, 77)

# only setting the value of c
display(c = 77)

rescale <- function(v, lower = 0, upper = 1) {
  # Rescales a vector, v, to lie in the range lower to upper.
  L <- min(v)
  H <- max(v)
  result <- (v - L) / (H - L) * (upper - lower) + lower
  return(result)
}



```

2. Create a function called `bootstrap()` that modifies the behaviour of the `sample()` function to perform a bootstrap sample on a given vector. Remember that bootstrap samples are the same **length** as the original sample, and remember that these samples are taken **with replacement** (*hint, hint*).
```{r}
bootstrap <- function(vector) {
  boot <- sample(vector, length(vector), replace = TRUE)
  return(boot)         
  }
```
3. Create the following vector: `grades = c(92, 80, 85, 95, 97, 78, 90, 86, 88, 66, 70, 82)`. Find the five-number summary, mean, and standard deviation of `grades`.
```{r}
grades <- c(92, 80, 85, 95, 97, 78, 90, 86, 88, 66, 70, 82)
summary(grades)
mean(grades)
sd(grades)
```

3. Call your function 10 times on `grades`. Store each bootstrapped sample as `boot1`, `boot2`, ..., `boot10`.
```{r}
boot1 <- c(bootstrap(grades))
boot2 <- c(bootstrap(grades))
boot3 <- c(bootstrap(grades))
boot4 <- c(bootstrap(grades))
boot5 <- c(bootstrap(grades))
boot6 <- c(bootstrap(grades))
boot7 <- c(bootstrap(grades))
boot8 <- c(bootstrap(grades))
boot9 <- c(bootstrap(grades))
boot10 <- c(bootstrap(grades))
```

4. Find the five-number summary, mean, and standard deviation of each bootstrap sample.
Boot1:
```{r}
summary(boot1)
mean(boot1)
sd(boot1)
```

Boot2:
```{r}
summary(boot2)
mean(boot2)
sd(boot2)
```

Boot3:
```{r}
summary(boot3)
mean(boot3)
sd(boot3)
```

Boot4:
```{r}
summary(boot4)
mean(boot4)
sd(boot4)
```

Boot5:
```{r}
summary(boot5)
mean(boot5)
sd(boot5)
```

Boot6:
```{r}
summary(boot6)
mean(boot6)
sd(boot6)
```

Boot7:
```{r}
summary(boot7)
mean(boot7)
sd(boot7)
```

Boot8:
```{r}
summary(boot8)
mean(boot8)
sd(boot8)
```

Boot9:
```{r}
summary(boot9)
mean(boot9)
sd(boot9)
```

Boot10:
```{r}
summary(boot10)
mean(boot10)
sd(boot10)
```

5. Create a data frame of these summary statistics (10 rows and seven columns).
```{r}
allBoots <- data.frame(
      BootID = c("boot1", "boot2", "boot3", "boot4", "boot5", "boot6", "boot7", "boot8", "boot9", "boot10"), 
      Minimum = c(summary(boot1)[1], summary(boot2)[1], summary(boot3)[1], summary(boot4)[1],  summary(boot5)[1], summary(boot6)[1], summary(boot7)[1], summary(boot8)[1], summary(boot9)[1], summary(boot10)[1]),
      FirstQuantile = c(summary(boot1)[2], summary(boot2)[2], summary(boot3)[2], summary(boot4)[2],  summary(boot5)[2], summary(boot6)[2], summary(boot7)[2], summary(boot8)[2], summary(boot9)[2], summary(boot10)[2]),
      Median = c(summary(boot1)[3], summary(boot2)[3], summary(boot3)[3], summary(boot4)[3],  summary(boot5)[3], summary(boot6)[3], summary(boot7)[3], summary(boot8)[3], summary(boot9)[3], summary(boot10)[3]),
      Mean = c(summary(boot1)[4], summary(boot2)[4], summary(boot3)[4], summary(boot4)[4],  summary(boot5)[4], summary(boot6)[4], summary(boot7)[4], summary(boot8)[4], summary(boot9)[4], summary(boot10)[4]),
      ThirdQuantile = c(summary(boot1)[5], summary(boot2)[5], summary(boot3)[5], summary(boot4)[5],  summary(boot5)[5], summary(boot6)[5], summary(boot7)[5], summary(boot8)[5], summary(boot9)[5], summary(boot10)[5]),  
      Maximum = c(summary(boot1)[6], summary(boot2)[6], summary(boot3)[6], summary(boot4)[6],  summary(boot5)[6], summary(boot6)[6], summary(boot7)[6], summary(boot8)[6], summary(boot9)[6], summary(boot10)[6]),   
      StandardDeviation = c(sd(boot1), sd(boot2), sd(boot3), sd(boot4), sd(boot5), sd(boot6), sd(boot7), sd(boot8), sd(boot9), sd(boot10)),
  stringsAsFactors = FALSE
)
```  

6. Find the bootstrap estimate of each element of the five-number summary, mean, and standard deviation (look at the formula in your notes, then look at the `?colMeans` help file).
```{r}
bootstrapMeanOfMin <- colMeans(allBoots[2])
bootstrapMeanOfFirstQ <- colMeans(allBoots[3])
bootstrapMeanOfMedian <- colMeans(allBoots[4])
bootstrapMeanOfMean <- colMeans(allBoots[5])
bootstrapMeanOfThirdQ <- colMeans(allBoots[6])
bootstrapMeanOfMax <- colMeans(allBoots[7])
bootstrapMeanOfSD <- colMeans(allBoots[8])

nminus1Boots <- 9

  bootMin1 <- ((allBoots[c(1), c(2)]-bootstrapMeanOfMin)^2)/nminus1Boots
  bootMin2 <- ((allBoots[c(2), c(2)]-bootstrapMeanOfMin)^2)/nminus1Boots
  bootMin3 <- ((allBoots[c(3), c(2)]-bootstrapMeanOfMin)^2)/nminus1Boots
  bootMin4 <- ((allBoots[c(4), c(2)]-bootstrapMeanOfMin)^2)/nminus1Boots
  bootMin5 <- ((allBoots[c(5), c(2)]-bootstrapMeanOfMin)^2)/nminus1Boots
  bootMin6 <- ((allBoots[c(6), c(2)]-bootstrapMeanOfMin)^2)/nminus1Boots
  bootMin7 <- ((allBoots[c(7), c(2)]-bootstrapMeanOfMin)^2)/nminus1Boots
  bootMin8 <- ((allBoots[c(8), c(2)]-bootstrapMeanOfMin)^2)/nminus1Boots
  bootMin9 <- ((allBoots[c(9), c(2)]-bootstrapMeanOfMin)^2)/nminus1Boots
  bootMin10 <- ((allBoots[c(10), c(2)]-bootstrapMeanOfMin))^2/nminus1Boots
  bootstrapMinSE <- bootMin1 + bootMin2 + bootMin3 + bootMin4 + bootMin5 + bootMin6 + bootMin7 + bootMin8 + bootMin9 + bootMin10
  
  bootFirstQ1 <- ((allBoots[c(1), c(3)]-bootstrapMeanOfFirstQ)^2)/nminus1Boots
  bootFirstQ2 <- ((allBoots[c(2), c(3)]-bootstrapMeanOfFirstQ)^2)/nminus1Boots
  bootFirstQ3 <- ((allBoots[c(3), c(3)]-bootstrapMeanOfFirstQ)^2)/nminus1Boots
  bootFirstQ4 <- ((allBoots[c(4), c(3)]-bootstrapMeanOfFirstQ)^2)/nminus1Boots
  bootFirstQ5 <- ((allBoots[c(5), c(3)]-bootstrapMeanOfFirstQ)^2)/nminus1Boots
  bootFirstQ6 <- ((allBoots[c(6), c(3)]-bootstrapMeanOfFirstQ)^2)/nminus1Boots
  bootFirstQ7 <- ((allBoots[c(7), c(3)]-bootstrapMeanOfFirstQ)^2)/nminus1Boots
  bootFirstQ8 <- ((allBoots[c(8), c(3)]-bootstrapMeanOfFirstQ)^2)/nminus1Boots
  bootFirstQ9 <- ((allBoots[c(9), c(3)]-bootstrapMeanOfFirstQ)^2)/nminus1Boots
  bootFirstQ10 <- ((allBoots[c(10), c(3)]-bootstrapMeanOfFirstQ))^2/nminus1Boots
  bootstrapFirstQSE <- bootFirstQ1 + bootFirstQ2 + bootFirstQ3 + bootFirstQ4 + bootFirstQ5 + bootFirstQ6 + bootFirstQ7 + bootFirstQ8 + bootFirstQ9 + bootFirstQ10

  bootMedian1 <- ((allBoots[c(1), c(4)]-bootstrapMeanOfMedian)^2)/nminus1Boots
  bootMedian2 <- ((allBoots[c(2), c(4)]-bootstrapMeanOfMedian)^2)/nminus1Boots
  bootMedian3 <- ((allBoots[c(3), c(4)]-bootstrapMeanOfMedian)^2)/nminus1Boots
  bootMedian4 <- ((allBoots[c(4), c(4)]-bootstrapMeanOfMedian)^2)/nminus1Boots
  bootMedian5 <- ((allBoots[c(5), c(4)]-bootstrapMeanOfMedian)^2)/nminus1Boots
  bootMedian6 <- ((allBoots[c(6), c(4)]-bootstrapMeanOfMedian)^2)/nminus1Boots
  bootMedian7 <- ((allBoots[c(7), c(4)]-bootstrapMeanOfMedian)^2)/nminus1Boots
  bootMedian8 <- ((allBoots[c(8), c(4)]-bootstrapMeanOfMedian)^2)/nminus1Boots
  bootMedian9 <- ((allBoots[c(9), c(4)]-bootstrapMeanOfMedian)^2)/nminus1Boots
  bootMedian10 <- ((allBoots[c(10), c(4)]-bootstrapMeanOfMedian))^2/nminus1Boots
  bootstrapMedianSE <- bootMedian1 + bootMedian2 + bootMedian3 + bootMedian4 + bootMedian5 + bootMedian6 + bootMedian7 + bootMedian8 + bootMedian9 + bootMedian10
  
  bootMean1 <- ((allBoots[c(1), c(5)]-bootstrapMeanOfMean)^2)/nminus1Boots
  bootMean2 <- ((allBoots[c(2), c(5)]-bootstrapMeanOfMean)^2)/nminus1Boots
  bootMean3 <- ((allBoots[c(3), c(5)]-bootstrapMeanOfMean)^2)/nminus1Boots
  bootMean4 <- ((allBoots[c(4), c(5)]-bootstrapMeanOfMean)^2)/nminus1Boots
  bootMean5 <- ((allBoots[c(5), c(5)]-bootstrapMeanOfMean)^2)/nminus1Boots
  bootMean6 <- ((allBoots[c(6), c(5)]-bootstrapMeanOfMean)^2)/nminus1Boots
  bootMean7 <- ((allBoots[c(7), c(5)]-bootstrapMeanOfMean)^2)/nminus1Boots
  bootMean8 <- ((allBoots[c(8), c(5)]-bootstrapMeanOfMean)^2)/nminus1Boots
  bootMean9 <- ((allBoots[c(9), c(5)]-bootstrapMeanOfMean)^2)/nminus1Boots
  bootMean10 <- ((allBoots[c(10), c(5)]-bootstrapMeanOfMean))^2/nminus1Boots
  bootstrapMeanSE <- bootMean1 + bootMean2 + bootMean3 + bootMean4 + bootMean5 + bootMean6 + bootMean7 + bootMean8 + bootMean9 + bootMean10

  bootThirdQ1 <- ((allBoots[c(1), c(6)]-bootstrapMeanOfThirdQ)^2)/nminus1Boots
  bootThirdQ2 <- ((allBoots[c(2), c(6)]-bootstrapMeanOfThirdQ)^2)/nminus1Boots
  bootThirdQ3 <- ((allBoots[c(3), c(6)]-bootstrapMeanOfThirdQ)^2)/nminus1Boots
  bootThirdQ4 <- ((allBoots[c(4), c(6)]-bootstrapMeanOfThirdQ)^2)/nminus1Boots
  bootThirdQ5 <- ((allBoots[c(5), c(6)]-bootstrapMeanOfThirdQ)^2)/nminus1Boots
  bootThirdQ6 <- ((allBoots[c(6), c(6)]-bootstrapMeanOfThirdQ)^2)/nminus1Boots
  bootThirdQ7 <- ((allBoots[c(7), c(6)]-bootstrapMeanOfThirdQ)^2)/nminus1Boots
  bootThirdQ8 <- ((allBoots[c(8), c(6)]-bootstrapMeanOfThirdQ)^2)/nminus1Boots
  bootThirdQ9 <- ((allBoots[c(9), c(6)]-bootstrapMeanOfThirdQ)^2)/nminus1Boots
  bootThirdQ10 <- ((allBoots[c(10), c(6)]-bootstrapMeanOfThirdQ))^2/nminus1Boots
  bootstrapThirdQSE <- bootThirdQ1 + bootThirdQ2 + bootThirdQ3 + bootThirdQ4 + bootThirdQ5 + bootThirdQ6 + bootThirdQ7 + bootThirdQ8 + bootThirdQ9 + bootThirdQ10
  
  bootMax1 <- ((allBoots[c(1), c(7)]-bootstrapMeanOfMax)^2)/nminus1Boots
  bootMax2 <- ((allBoots[c(2), c(7)]-bootstrapMeanOfMax)^2)/nminus1Boots
  bootMax3 <- ((allBoots[c(3), c(7)]-bootstrapMeanOfMax)^2)/nminus1Boots
  bootMax4 <- ((allBoots[c(4), c(7)]-bootstrapMeanOfMax)^2)/nminus1Boots
  bootMax5 <- ((allBoots[c(5), c(7)]-bootstrapMeanOfMax)^2)/nminus1Boots
  bootMax6 <- ((allBoots[c(6), c(7)]-bootstrapMeanOfMax)^2)/nminus1Boots
  bootMax7 <- ((allBoots[c(7), c(7)]-bootstrapMeanOfMax)^2)/nminus1Boots
  bootMax8 <- ((allBoots[c(8), c(7)]-bootstrapMeanOfMax)^2)/nminus1Boots
  bootMax9 <- ((allBoots[c(9), c(7)]-bootstrapMeanOfMax)^2)/nminus1Boots
  bootMax10 <- ((allBoots[c(10), c(7)]-bootstrapMeanOfMax))^2/nminus1Boots
  bootstrapMaxSE <- bootMax1 + bootMax2 + bootMax3 + bootMax4 + bootMax5 + bootMax6 + bootMax7 + bootMax8 + bootMax9 + bootMax10  
  
  bootSD1 <- ((allBoots[c(1), c(7)]-bootstrapMeanOfSD)^2)/nminus1Boots
  bootSD2 <- ((allBoots[c(2), c(7)]-bootstrapMeanOfSD)^2)/nminus1Boots
  bootSD3 <- ((allBoots[c(3), c(7)]-bootstrapMeanOfSD)^2)/nminus1Boots
  bootSD4 <- ((allBoots[c(4), c(7)]-bootstrapMeanOfSD)^2)/nminus1Boots
  bootSD5 <- ((allBoots[c(5), c(7)]-bootstrapMeanOfSD)^2)/nminus1Boots
  bootSD6 <- ((allBoots[c(6), c(7)]-bootstrapMeanOfSD)^2)/nminus1Boots
  bootSD7 <- ((allBoots[c(7), c(7)]-bootstrapMeanOfSD)^2)/nminus1Boots
  bootSD8 <- ((allBoots[c(8), c(7)]-bootstrapMeanOfSD)^2)/nminus1Boots
  bootSD9 <- ((allBoots[c(9), c(7)]-bootstrapMeanOfSD)^2)/nminus1Boots
  bootSD10 <- ((allBoots[c(10), c(7)]-bootstrapMeanOfSD))^2/nminus1Boots
  bootstrapSDSE <- bootSD1 + bootSD2 + bootSD3 + bootSD4 + bootSD5 + bootSD6 + bootSD7 + bootSD8 + bootSD9 + bootSD10  
```
